---
title: "Graphs"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 100)
```

# Comparison of diverse graph clustering algorithms.

```{r, message=FALSE}
library(igraph)
dat <- read.table("graph_ex1.txt", header = FALSE, sep = " ")
mis_graph <- simplify(graph.data.frame(dat, directed=FALSE))
```

* Size and order of the graph

```{r, message=FALSE}
size = ecount(mis_graph)
order = vcount(mis_graph)
c(size, order)
```

* Density and diameter of the graph

```{r, message=FALSE}
diam = diameter(mis_graph)
diam 
```

* Mean distance

```{r, message=FALSE}
md = mean_distance(mis_graph, directed = FALSE)
md
```

* Plot the graph
```{r, fig.height = 10, fig.width = 10}
set.seed(3)
V(mis_graph)$label.cex <- (degree(mis_graph)+10)/max(degree(mis_graph)) * 1.5
l <- layout.fruchterman.reingold(mis_graph)
plot(mis_graph, vertex.size = 0, layout = l)
```

## Edge betweenness##


```{r, fig.height = 10, fig.width = 10} 
set.seed(3)
# Edge betweenness 
eb <- cluster_edge_betweenness(mis_graph, directed=F)
dendPlot(eb, mode="hclust")
```

### Clustering ###

## Hierarchical cluster analysis (Jaccard similarity)

```{r, fig.height=10, fig.width=10, message=FALSE}
d <- 1-similarity(mis_graph, method = 'jaccard')
d <- as.dist(d)
res <- hclust(d, method="complete")
# Plot the dendrogram
plot(res, labels=V(mis_graph)$name, hang=-1, cex=0.8)
``` 

```{r, fig.height = 10, fig.width = 10}
lab_hclust = cutree(res, 3) # associated clustering (e.g. 3 clusters)
plot(mis_graph, layout = l, vertex.size = 5, vertex.color = lab_hclust)
```

## Edge betweenness

```{r, fig.height = 10, fig.width = 10} 
# Edge betweenness 
plot(mis_graph, layout = l, vertex.size = 5, vertex.color = eb$membership)
```

## Louvain algorithm ##

```{r, fig.height = 10, fig.width = 10}
# Louvain 
louv <- cluster_louvain(mis_graph)
plot(mis_graph, layout = l, vertex.size = 5, vertex.color = louv$membership)
```
## Spectral clustering ##

Degrees and indicence matrix
Normalized graph Laplacian
```{r, fig.height = 10, fig.width = 10}
# library(network)
# library(intergraph)
D <- diag(degree(mis_graph))
SD <- diag(degree(mis_graph)**(-0.5))
A <- get.adjacency(mis_graph)
L <- D - A
SL <- SD%*%L%*%SD
EL <- eigen(SL)
VL <- EL$vectors
VL <- data.frame(VL)
row.names(VL) <- attributes(A)$Dimnames[[1]]
i = 8
j = 9
plot(VL[,i], VL[,j], type="n", xlab=paste("Axis",i), ylab=paste("Axis",j))
text(VL[,i], VL[,j], row.names(VL))
```

### Unnormalized graph Laplacian
```{r, fig.height = 10, fig.width = 10}
E <- eigen(L)
V <- E$vectors
V <- data.frame(V)
row.names(V) <- attributes(A)$Dimnames[[1]]
i = 8
j = 9
plot(V[,i], V[,j], type="n", xlab=paste("Axis",i), ylab=paste("Axis",j))
text(V[,i], V[,j], row.names(V))
```

```{r}
D
A
L
```

### Using clustering_leading_eigen in igraphs
```{r, fig.height = 10, fig.width = 10}
# SC
sc = cluster_leading_eigen(mis_graph)
V(mis_graph)$color = sc$membership

plot.igraph(mis_graph, vertex.size = 5, vertex.label=V(mis_graph)$name, layout=l, edge.color="black", edge.width=E(mis_graph)$weight/5000)

```
