---
title: "Exercise 2"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 200)
```
Let us consider 40 kinds of beers, each one being characterized by the four
following variables:  

Variable | Meaning  
---------|------------------------------------------  
Taste    | Quantitative appreciation of taste given by experts  
Bitter   | High values correspond to bitter beers, low values to sweet beers  
Thirst   | High values correspond to thirst-quenching beers, low values to beers that leave consumers still thirsty  
DgAlcohol | Degree of alcohol  

The aim of the study is to use principal component analysis (PCA) to 
obtain a summary of the data and visualize it.

## Question 1: ##
The result of PCA is loaded by
```{r}
load("beers.pca")
```
Provide and execute ```R``` code to plot individuals in the first principal plane.  

### Answer: ###

## Question 2: ##

Provide and execute ```R``` code to perform PCA on variables.  
Plot variables in the first principal plane. 
Three groups of variables clearly appear in this figure.  
What can be said about correlations:    

* Between variables belonging to different groups?
* Between variables belonging to the same group?

### Answer: ###

## Question 3: ##

In PCA computations, was the ```scale``` argument set to ```FALSE``` or ```TRUE```?  
How can you tell?  
For which purpose was ```scale``` set to that value?

### Answer: ###

## Question 4: ##

Do you assess that two principal components are sufficient to represent the features of the beers?   
Why?  
If not, how many principal components would be needed?  

### Answer: ###

## Question 5: ##

Now return to the figure provided in question 1.  
For each of the four quadrants (upper-left, upper-right, 
and so on): what properties are shared by the beers within this quadrant?  
On average, drinking more than 50cl beer results into 
the legal inability to drive a car. You wish to drink 50cl beer and
quench your thirst, but you wish also to minimise the risk of having
a fee for exceeding the legal rate of alcohol while driving your car.
Which beer should you avoid in this case? (give detailed justification).  
Which beer would you choose if you wanted to avoid bitterness?  
What potential drawback of this beer should you expect
(give detailed justification)?  
Which beer would you choose if you would like to avoid both bitterness 
and bad taste (give detailed justification)?  

### Answer: ###