---
title: "Exercise 1"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 200)
```
Let consider the following R script:
```{r}
set.seed(0)
# Number of variables
pt = 201
# Number of predictors
p = pt - 1
# Sample size
n = 30 * p
D = matrix(rnorm(n*pt), nrow=n, ncol=pt)
D = data.frame(D)
names(D)[pt] = "Y"
reg = lm(Y~., data=D)
```
## Question 1: ##
* What does the script above do?
* What is the true distribution for the random variable $Y$ given the first 200 columns of ```D```: $X_1, \ldots, X_200$ (referred to as “variables”). Give the name and parameters of this distribution.

### Answer: ###

## Question 2: ##
* Write an equation defining model estimated by the ```lm``` command in line 21. What is the difference between this model and the one defined in question 1?
* Hint: write equations using 
<pre><code>
$$
(a+b)^n = \sum\limits_{k=0}^n \frac{n!}{k! (n-k)!} a^k b^{n-k}.
$$
</code></pre>

### Answer: ###

## Question 3: ##
Provide and execute some ```R``` code to compute the number of coefficient assessed as significantly non-zero at level 5%.

Hint: ```summary(reg)$coefficients```

### Answer: ###

## Question 4: ##
Provide an explanation for the result obtained in question 3.

### Answer: ###

## Question 5: ##
What issue is raised by the result obtained in question 3?

### Answer: ###

## Question 6: ##
Describe a possible solution to solve this problem, explaining why you think it should work (with full details and defining the concepts you introduce: 10 lines expected, assuming approx. 85 characters / line in your editor).

Implement that solution and comment the results.

### Answer: ###

