---
title: "Exercise 1: regression"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 200)
```
Car consumptions are measured in MPGs (miles per gallon): higher MPG values mean that the car is able to do more miles with one gallon of fuel.

To determine the car features having an effect on consumption, we use the following linear regression model.
```{r}
M <- lm(mpg~., data=mtcars)
summary(M)
```

## Question 1: ##

Which predictors seem to have an effect on mpg? 

Make your criterion explicit.

### Answer: ###

## Question 2: ##

Provide a 95% confidence interval for the ```carb``` parameter. 

What can you deduce from the fact that 0 is in the confidence interval?

Provide descriptions and formulas to explain what the quantity in column 
```Pr(>|t|)``` represents.

### Answer: ### 

## Question 3: ##

Provide descriptions and formulas to explain what ```p-value: 3.793e-07``` represents. 

What to conclude from this value?

Does it seem in discordance with your answer in question 1? In this case, what could explain such discordance?

### Answer: ### 


## Question 4: ##

What strategy could be used to obtain some linear regression model with a maximal number of predictors, all of which having significant effect at level 5%? Can you run it in practice? If yes, provide the estimates. If not, provide the reason for it.

### Answer: ### 


## Question 5: ##

Describe and run some algorithm to find the best model in some sense.


### Answer: ### 
